---
title: "MA615 Final Project"
author: "Kaiwei Xiao"
date: "2022-12-17"
output:
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE,results='hide'}
library(ggplot2)
library(dplyr)
library(svglite)
library(devtools)
library(forecast)
library(tseries)
library(ggplot2)
library(ggfortify)
library(fpp)
library(vars)
```

```{r}
#2022 Quarter 1 Green lines travel data 
Q1L = read.csv("2022-Q1_LRTravelTimes.csv")
```

```{r}
##2022 Quarter 1 other lines travel data 
Q1H = read.csv("2022-Q1_HRTravelTimes.csv")

Q2L = read.csv("2022-Q2_LRTravelTimes.csv")

Q2H = read.csv("2022-Q2_HRTravelTimes.csv")

Q3L = read.csv("2022-Q3_LRTravelTimes.csv")

Q3H = read.csv("2022-Q3_HRTravelTimes.csv")
```

```{r}
# Append the quarter 1 dataset
Q1 = rbind(Q1H, Q1L)
head(Q1)
```

```{r}
#check how many lines in this dataset
levels(factor(Q1$route_id))
```

There are couple lines here.Green(including subset), Blue, Orange and
Red lines. There is also a line called "Mattapan".

```{r}
#compare average travel time of different lines
travelt = aggregate(travel_time_sec ~ route_id, data = Q1, mean)
travelt$travel_time_sec = round(travelt$travel_time_sec, digits = 1)
travelt
```

```{r}
ggplot(data = travelt, aes(x = route_id, y = travel_time_sec, fill = route_id)) + 
  geom_bar(stat = "identity")+
  geom_text(aes(label= travel_time_sec), vjust=1.6, color="black", size=3.5)+
  theme_minimal()
```

As we can see, the Mattapan has the shortest average travel time. The
Green-D line has the longest average travel time which is 1005.7
seconds. It is almost 16.7 minutes for it to travel from a stop to the
next one! ![Greenline D is so
long](/Users/j/MA615/final%20project/GreenlineD.jpg) It is very
reasonable because as you can see, the Green line D is a long line, and
the distance from stop to stop is quite long.

```{r}
#Let's see how travel time impacted by the date.
traveld = aggregate(travel_time_sec ~ service_date, data = Q1, mean)
traveld$travel_time_sec = round(traveld$travel_time_sec, digits = 1)
head(traveld)
```

```{r}
#Plot average travel time vs date
p = ggplot(traveld, aes(x=as.Date(service_date), y=travel_time_sec)) +
  geom_line() 
p + scale_x_date(date_minor_breaks = "2 day")
```

```{r}
#see which date got the longest mean travel time.
traveld[which.max(traveld$travel_time_sec),]
```

```{r}
#see which date got the shortest mean travel time.
traveld[which.min(traveld$travel_time_sec),]
```

It seems that in the quarter 1 time range, the travel time trend is not
that obvious. How about looking through the 2022 year?

```{r}
#combine the 2022
QA = rbind(Q1H, Q1L, Q2H, Q2L, Q3H, Q3L)
head(QA)
```

```{r}
#let's see if there is any new lines appear.
levels(factor(QA$route_id))
```

Still, we got the same line in the rest of the year.

```{r}
#let's check the travel time by date through 2022
travelda = aggregate(travel_time_sec ~ service_date, data = QA, mean)
travelda$travel_time_sec = round(travelda$travel_time_sec, digits = 1)
head(travelda)
```

```{r}
#Plot average travel time vs date
pa = ggplot(travelda, aes(x=as.Date(service_date), y=travel_time_sec)) +
  geom_line() 
pa + scale_x_date(date_minor_breaks = "2 day")
```

```{r}
#see which date got the shortest mean travel time.
travelda[which.min(travelda$travel_time_sec),]
#see which date got the longest mean travel time.
travelda[which.max(travelda$travel_time_sec),]
```

The shortest mean travel time happened on March 13,2022 and it was 674.2
sec. The longest mean travel time happened on September 19,2022 and it
was 894 sec.

Can we build a time series model for the travel time?

```{r}
pacf(travelda$travel_time_sec)
```

```{r}
#take difference for time seiries model building
dif_travel = diff(travelda$travel_time_sec)
```

```{r}
#plot the time series data
autoplot(
  pacf(dif_travel, lag.max = 30), 
  main="corellogram of time series"
)
```

As an inital pass, we will fit ARIMA(1,1,1) with seasonal difference.

```{r}
model_arima <- Arima(
  y = travelda$travel_time_sec, 
  order = c(1, 1, 1),
  seasonal = list(order = c(0, 1, 0))
)

model_arima
```

```{r}
checkresiduals(model_arima)
```

The Ljung-Box test test the overall randomness of the time seiries. H0:
The data are independently distributed (i.e. the correlations in the
population from which the sample is taken are 0, so that any observed
correlations in the data result from randomness of the sampling
process). Ha: The data are not independently distributed; they exhibit
serial correlation.

Here the test p-value is significant, we can reject the null and
conclude that the data are not independently distributed, there is
serial correlation here.

Now we use auto.arima to see what we get using Hyndman & Khandakar
(2008) algorithm. This algorithm can help us choose the p and q for the
ARIMA model based on AIC.

```{r}
sarimax_petro_law = auto.arima(
  y = travelda$travel_time_sec, 
  ic = "aic",
  max.order = 10,
  stepwise = F,
  approximation = F,
  parallel = T,
  num.cores = 4
)

# best model
sarimax_petro_law

```

It choose p = 4 and q = 5 with no difference(I=0) for the model
automatically. Also we can see the all year mean travel time for all
lines is 814.55 sec.

```{r}
checkresiduals(sarimax_petro_law)
```

The Ljung-Box test tell us there is serial correlation here and the
model is meaningful.

We can write the model in equation here: Average_travel_time =
814.6+1.17Y(t-1)-0.69Y(t-2)+0.06Y(t-3)+0.31Y(t-4)
-0.56e(t-1)+0.28e(t-2)+0.24e(t-3)-0.31e(t-4)-0.17(t-5)+error

The Y(t) means the p days average travel time in front of the date you
want to predict, and the e(t) means the white noise of series in front
of the date you want to predict and the error is the error!

So if you want to predict the average travel time by date, use this
model.

##Shiny app

for the shiny app there are two version

The first version is driving map, which allow you to add waypoint on it.
Simply type in the departure address and final destination (waypoints if
needed), you can get the driving routine. Also use your map key and API
key if possible.

#published link: <https://jpz61m-jeremy-x0.shinyapps.io/Driving_Map/>

The second version is transit map. It allows you travel through public
transportation but no waypoints due to the API limitation.

#published link: <https://jpz61m-jeremy-x0.shinyapps.io/Transit_Map/>
